---
title: "Pilot_17_OOO_Analysis"
output: html_document
date: "2025-12-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE}
library(readr)
library(tidyverse)
library(stringr)
library(survival)
library(lme4)
library(tidyr)

library(broom)
library(dplyr)
library(ggplot2)

library(vegan)
```

```{r, echo = FALSE}
df1 <- read_csv("Han_curvRep1D_135373_SESSION_2025-12-05_15h44.11.905.csv")
df2 <- read_csv("Han_curvRep1D_202295_SESSION_2025-12-03_23h17.34.494.csv")
df3 <- read_csv("Han_curvRep1D_241598_SESSION_2025-12-05_23h58.47.45.csv")
df4 <- read_csv("Han_curvRep1D_273589_SESSION_2025-12-04_16h19.53.591.csv")
df5 <- read_csv("Han_curvRep1D_346311_SESSION_2025-12-05_21h47.10.211.csv")
df6 <- read_csv("Han_curvRep1D_380418_SESSION_2025-12-05_22h47.42.192.csv")
df7 <- read_csv("Han_curvRep1D_411219_SESSION_2025-12-05_03h29.35.845.csv")
df8 <- read_csv("Han_curvRep1D_513908_SESSION_2025-12-02_12h24.20.787.csv")
df9 <- read_csv("Han_curvRep1D_645088_SESSION_2025-12-05_21h04.09.161.csv")
df10 <- read_csv("Han_curvRep1D_712067_SESSION_2025-12-01_20h26.13.788.csv")
df11 <- read_csv("Han_curvRep1D_720702_SESSION_2025-12-03_12h06.17.516.csv")
df12 <- read_csv("Han_curvRep1D_770450_SESSION_2025-12-02_18h03.42.69.csv")
df13 <- read_csv("Han_curvRep1D_797896_SESSION_2025-12-04_18h55.19.676.csv")
df14 <- read_csv("Han_curvRep1D_879809_SESSION_2025-12-04_23h05.25.59.csv")
df15 <- read_csv("Han_curvRep1D_909758_SESSION_2025-12-04_21h23.15.701.csv")
df16 <- read_csv("Han_curvRep1D_913438_SESSION_2025-12-05_20h58.37.81.csv")
df17 <- read_csv("Han_curvRep1D_967429_SESSION_2025-11-27_22h18.59.884.csv")

```



```{r}
library(readr)
library(tidyverse)

# filenames in a vector
files <- c(
  "Han_curvRep1D_135373_SESSION_2025-12-05_15h44.11.905.csv",
  "Han_curvRep1D_202295_SESSION_2025-12-03_23h17.34.494.csv",
  "Han_curvRep1D_241598_SESSION_2025-12-05_23h58.47.45.csv",
  "Han_curvRep1D_273589_SESSION_2025-12-04_16h19.53.591.csv",
  "Han_curvRep1D_346311_SESSION_2025-12-05_21h47.10.211.csv",
  "Han_curvRep1D_380418_SESSION_2025-12-05_22h47.42.192.csv",
  "Han_curvRep1D_411219_SESSION_2025-12-05_03h29.35.845.csv",
  "Han_curvRep1D_513908_SESSION_2025-12-02_12h24.20.787.csv",
  "Han_curvRep1D_645088_SESSION_2025-12-05_21h04.09.161.csv",
  "Han_curvRep1D_712067_SESSION_2025-12-01_20h26.13.788.csv",
  "Han_curvRep1D_720702_SESSION_2025-12-03_12h06.17.516.csv",
  "Han_curvRep1D_770450_SESSION_2025-12-02_18h03.42.69.csv",
  "Han_curvRep1D_797896_SESSION_2025-12-04_18h55.19.676.csv",
  "Han_curvRep1D_879809_SESSION_2025-12-04_23h05.25.59.csv",
  "Han_curvRep1D_909758_SESSION_2025-12-04_21h23.15.701.csv",
  "Han_curvRep1D_913438_SESSION_2025-12-05_20h58.37.81.csv",
  "Han_curvRep1D_967429_SESSION_2025-11-27_22h18.59.884.csv"
)

df_raw_all <- map_dfr(files, function(f) {
  read_csv(f, 
           col_types = cols(rt = col_double(), 
                            response = col_character(), 
                            correct_response = col_character(), 
                            correct = col_logical(), 
                            block_failed = col_logical(), 
                            blocks_failed_total = col_double(), 
                            .default = col_guess()),
           na = c("", "NA", "null", "NaN")) %>%
    mutate(
      file = basename(f),
      subj_id = str_match(file, "Han_curvRep1D_(\\d+)_SESSION")[,2]
    )
})

```


```{r}
df_raw_all %>% count(subj_id, trial_kind) %>% arrange(subj_id, desc(n)) %>% filter(trial_kind == "exp_221" | trial_kind == "exp_111")

# in experiment design code, when generating all the possible triplets of the patterns 2/2/2 and 2/1/1, there should have been 324 exp_222 and 216 exp_111. so 540 total, and all these triplets should be used. All participants see the same 540 triplets, since these are all the combinations possible with those patterns. 
# But instead of 216 exp_111, there were 270 total. (54 more that should not be included, that were the 111 triplets that hold one dimension fixed). 
# so when we subsample 540 of them after shuffling, it takes the first 540 of the 594 total possible triplets. this means that the number of exp_221 and exp_111 trials are not constant between participants. 
# participants don't all see all the same triplets 
```

```{r}
# catch trial accuracy for practice trials
df_raw_all %>%
  filter(trial_kind == "practice_catch") %>%
  group_by(subj_id) %>%
  summarise(
    n_practice_catch = n(),
    n_correct = sum(correct == TRUE, na.rm = TRUE),
    acc = mean(correct == TRUE, na.rm = TRUE)
  )

# catch trial accuracy for main trials
main_catch <- df_raw_all %>%
  filter(trial_kind == "catch")

main_catch %>%
  group_by(subj_id) %>%
  summarise(
    n_catch = n(),
    n_correct = sum(correct == TRUE, na.rm = TRUE),
    accuracy = mean(correct == TRUE, na.rm = TRUE))
```
```{r}
# each participant must get 80% catch trial accuracy min in each block to pass that block
block_catch_summary <- df_raw_all %>%
  filter(trial_kind == "catch") %>%
  group_by(subj_id, block_index) %>%
  summarise(
    n_catch = n(),
    n_correct = sum(correct == TRUE, na.rm = TRUE),
    acc = n_correct / n_catch,
    block_failed_80 = acc < 0.80,
    .groups = "drop"
  )
block_catch_summary %>% filter(block_failed_80 == TRUE) 

df_raw_all %>%
  filter(trial_kind == "block_feedback") %>%
  select(subj_id, block_index, block_failed, blocks_failed_total)

# no one failed any blocks 
```

```{r}
# Checking the RT warnings (= RTs faster than 500)
rt_warnings <- df_raw_all %>%
  filter(trial_kind == "rt_warning") %>%
  count(subj_id)
rt_warnings

df_raw_all %>%
  filter(rt < 500) %>% 
  select(participant, rt, block_index, trial_in_block, trial_kind)
```
```{r}
# clean the table. drop too fast trials and only keep exp221 and exp111
RT_MIN_MS <- 500
df_trials <- df_raw_all %>%
  filter(
    trial_kind %in% c("exp_221", "exp_111"),
    is_practice == FALSE,
    is_catch == FALSE,
    rt >= RT_MIN_MS
  )

# each row = one item in the triad. 
df_items <- df_trials %>%
  group_by(subj_id) %>%
  mutate(triad_id = row_number()) %>%
  ungroup() %>%
  select(
    subj_id, triad_id, chosen_index,
    stim_left, stim_middle, stim_right
  ) %>%
  pivot_longer(
    cols = c(stim_left, stim_middle, stim_right),
    names_to = "position",
    values_to = "stim_path"
  ) %>%
  mutate(
    pos_index = case_when(
      position == "stim_left"   ~ 0L,
      position == "stim_middle" ~ 1L,
      position == "stim_right"  ~ 2L
    ),
    is_chosen = pos_index == chosen_index,
    nv   = as.integer(str_match(stim_path, "stim_nv(\\d+)_")[,2]),
    ang  = as.integer(str_match(stim_path, "ang(\\d+)_curv")[,2]),
    curv = as.integer(str_match(stim_path, "curv(\\d+)_rot")[,2])
  )

head(df_items)
```



```{r}
library(dplyr)

items_short <- df_items %>%
  select(subj_id, triad_id, pos_index, is_chosen, curv, nv, ang)

pairs <- df_items %>%
  select(subj_id, triad_id, pos_index, is_chosen, curv, nv, ang) %>%
  inner_join(items_short, by = c("subj_id", "triad_id"), suffix = c("_i", "_j")) %>%
  filter(pos_index_i < pos_index_j) %>%
  arrange(subj_id, triad_id, pos_index_i, pos_index_j) %>%
  group_by(subj_id, triad_id) %>%
  mutate(
    pair_in_triad = row_number(),
    trial_id = cur_group_id() 
  ) %>%
  ungroup() %>%
  transmute(
    subj_id,
    triad_id,
    pair_in_triad,
    curv_i = curv_i, nv_i = nv_i, ang_i = ang_i,
    curv_j = curv_j, nv_j = nv_j, ang_j = ang_j,
    chosen_i = is_chosen_i,
    chosen_j = is_chosen_j,
    same_curv  = as.integer(curv_i == curv_j),
    same_vert  = as.integer(nv_i   == nv_j),
    same_angle = as.integer(ang_i  == ang_j),
    co_group   = as.integer(!chosen_i & !chosen_j),
    diff_curv  = as.factor(abs(curv_i - curv_j)),
    diff_nv    = as.factor(abs(nv_i   - nv_j)),
    diff_ang   = as.factor(abs(ang_i  - ang_j)),
    trial_id
  )

head(pairs)

# exactly 3 pairs per triad
pairs %>%
  count(subj_id, triad_id) %>%
  count(n)

# co_group exactly 1 per triad
pairs %>%
  group_by(subj_id, triad_id) %>%
  summarise(n_co = sum(co_group), .groups="drop") %>%
  count(n_co)

```
```{r}
library(lme4)

model1 <- glmer(
  co_group ~ same_curv + same_vert + same_angle +
    (1 | subj_id),
  data = pairs,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 2e5)
  )
)

summary(model1)
fixef(model1)        # log-odds
exp(fixef(model1))  # odds ratios


```
```{r}
contrasts(pairs$diff_curv) <- contr.sum(nlevels(pairs$diff_curv))
contrasts(pairs$diff_nv)   <- contr.sum(nlevels(pairs$diff_nv))
contrasts(pairs$diff_ang)  <- contr.sum(nlevels(pairs$diff_ang))

m_dist <- glmer(
  co_group ~ diff_curv + diff_nv + diff_ang +
    (1 | subj_id),
  data = pairs,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 2e5)
  )
)

summary(m_dist)

```
```{r}
contrasts(pairs$diff_curv)
contrasts(pairs$diff_nv)
contrasts(pairs$diff_ang)
```

```{r}
library(dplyr)
library(tidyr)

# 1) Start from df_items and add chosen as 0/1
df_long <- df_items %>%
  mutate(chosen = as.integer(is_chosen))

# 2) Build all within-triad pairs (i != j), compute pairwise diffs, then sum them per focal item i
pairs2 <- df_long %>%
  select(subj_id, triad_id, pos_index, nv, ang, curv) %>%
  rename(i = pos_index,
         nv_i = nv, ang_i = ang, curv_i = curv) %>%
  inner_join(
    df_long %>%
      select(subj_id, triad_id, pos_index, nv, ang, curv) %>%
      rename(j = pos_index,
             nv_j = nv, ang_j = ang, curv_j = curv),
    by = c("subj_id", "triad_id")
  ) %>%
  filter(i != j) %>%
  mutate(
    d_nv   = abs(nv_i   - nv_j),
    d_curv = abs(curv_i - curv_j),

    # circular angle distance for 180°-periodic orientations:
    # min(|a-b|, 180-|a-b|)
    d_ang  = pmin(abs(ang_i - ang_j), 180 - abs(ang_i - ang_j))
  ) %>%
  group_by(subj_id, triad_id, i) %>%
  summarise(
    diff_nv   = sum(d_nv),
    diff_curv = sum(d_curv),
    diff_ang  = sum(d_ang),
    .groups = "drop"
  )

df_long <- df_long %>%
  left_join(
    pairs2 %>% rename(pos_index = i),
    by = c("subj_id", "triad_id", "pos_index")
  )


library(nnet)

m_main <- multinom(chosen ~ diff_curv + diff_nv + diff_ang, data = df_long)

summary(m_main)

```

Multinomial Logistic 

```{r}
library(dplyr)
library(mlogit)

pairs2 <- pairs %>%
  mutate(
    chid   = interaction(subj_id, triad_id, drop = TRUE),  
    alt    = factor(pair_in_triad),
    choice = as.integer(co_group == 1)
  ) %>%
  distinct(chid, alt, .keep_all = TRUE) 

pairs2_df <- pairs2 %>%
  mutate(
    chid = as.character(chid),
    alt  = as.character(alt)
  ) %>%
  arrange(chid, alt)

pairs2_df <- as.data.frame(pairs2_df)

ml_dat <- mlogit.data(
  data     = pairs2_df,
  choice   = "choice",
  shape    = "long",
  chid.var = "chid",
  alt.var  = "alt",
  id.var   = "subj_id",
  panel    = TRUE
)

mA1_dist <- mlogit(choice ~ diff_curv + diff_nv + diff_ang | 0, data = ml_dat)
summary(mA1_dist)

```

Binary Logistic 

```{r}
library(lme4)

m_4 <- glmer(
  co_group ~ same_curv + same_vert + same_angle +
    (1 | subj_id) +
    (1 | subj_id:triad_id),
  data = pairs,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 2e5)
  )
)

m_5 <- glmer(
  co_group ~ diff_curv + diff_nv + diff_ang +
    (1 | subj_id) +
    (1 | subj_id:triad_id),
  data = pairs,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 2e5)
  )
)

summary(m_4)
summary(m_5)

```



RDM ANALYSIS 

```{r}
stim_lookup <- pairs %>%
  select(curv = curv_i, nv = nv_i, ang = ang_i) %>%
  bind_rows(pairs %>% select(curv = curv_j, nv = nv_j, ang = ang_j)) %>%
  distinct(curv, nv, ang) %>%
  arrange(curv, nv, ang) %>%
  mutate(stim_id = row_number())

stopifnot(nrow(stim_lookup) == 27)

pairs_id <- pairs %>%
  left_join(stim_lookup, by = c("curv_i"="curv", "nv_i"="nv", "ang_i"="ang")) %>%
  rename(id_i = stim_id) %>%
  left_join(stim_lookup, by = c("curv_j"="curv", "nv_j"="nv", "ang_j"="ang")) %>%
  rename(id_j = stim_id) %>%
  mutate(
    low  = pmin(id_i, id_j),
    high = pmax(id_i, id_j) # so that (i,j) and (j,i) become the same pair
  )

```

```{r}
# for each subject, S = n_win/n_appear
S_subj <- pairs_id %>%
  group_by(subj_id, low, high) %>%
  summarise(
    n_appear = n(),
    n_win    = sum(co_group),
    S        = n_win / n_appear,
    .groups = "drop"
  )

# average across 17 subjects
S_avg <- S_subj %>%
  group_by(low, high) %>%
  summarise(
    S_mean = mean(S, na.rm = TRUE),
    S_sd   = sd(S, na.rm = TRUE),
    S_se   = S_sd / sqrt(n()),
    n_subj = n(),
    .groups = "drop"
  )

```

```{r}
n_items <- 27

Sbar <- matrix(NA_real_, nrow = n_items, ncol = n_items)
rownames(Sbar) <- colnames(Sbar) <- as.character(1:n_items)

# fill off-diagonals
for (k in seq_len(nrow(S_avg))) {
  i <- S_avg$low[k]
  j <- S_avg$high[k]
  Sbar[i, j] <- S_avg$S_mean[k]
  Sbar[j, i] <- S_avg$S_mean[k]
}

diag(Sbar) <- 1

Dbar <- 1 - Sbar
diag(Dbar) <- 0
```

```{r}
S_avg %>% count(n_subj) %>% arrange(n_subj) # n_subj = 17 for all 243 pairs

```

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)

### SIMILARITY MATRIX HEATMAP

Sbar_long <- as_tibble(Sbar, .name_repair = "minimal") %>%
  mutate(id_i = row_number()) %>%
  pivot_longer(
    cols = -id_i,
    names_to = "id_j",
    values_to = "S"
  ) %>%
  mutate(
    id_i = as.integer(id_i),
    id_j = as.integer(id_j))

# make sure it's ordered by curv, nv, ang
stim_order <- stim_lookup %>%
  arrange(curv, nv, ang) %>%
  pull(stim_id)
Sbar_long <- Sbar_long %>%
  mutate(
    id_i = factor(id_i, levels = stim_order),
    id_j = factor(id_j, levels = stim_order))

ggplot(Sbar_long, aes(x = id_i, y = id_j, fill = S)) +
  geom_tile() +
  scale_fill_viridis_c(limits = c(0, 1), name = "Similarity", na.value = "grey90") +
  coord_equal() +
  theme_minimal(base_size = 14) +
  labs(
    x = "Stimulus",
    y = "Stimulus",
    title = "Average similarity matrix (S̄)"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )


### DISSIMILARITY MATRIX HEATMAP

Dbar_long <- as_tibble(Dbar, .name_repair = "minimal") %>%
  mutate(id_i = row_number()) %>%
  pivot_longer(
    cols = -id_i,
    names_to = "id_j",
    values_to = "D"
  ) %>%
  mutate(
    id_i = as.integer(id_i),
    id_j = as.integer(id_j)
  ) %>%
  mutate(
    id_i = factor(id_i, levels = stim_order),
    id_j = factor(id_j, levels = stim_order)
  )

ggplot(Dbar_long, aes(x = id_i, y = id_j, fill = D)) +
  geom_tile() +
  scale_fill_viridis_c(
    limits = c(0, 1),
    name = "Dissimilarity",
    na.value = "grey90"
  ) +
  coord_equal() +
  theme_minimal(base_size = 14) +
  labs(
    x = "Stimulus",
    y = "Stimulus",
    title = "Average dissimilarity matrix (D̄ = 1 − S̄)"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )

```

```{r}
# D̄ = 1 - S̄
Dbar <- 1 - Sbar
diag(Dbar) <- 0

Sbar_long <- as_tibble(Sbar, .name_repair = "minimal") %>%
  mutate(id_i = row_number()) %>%
  pivot_longer(cols = -id_i, names_to = "id_j", values_to = "val") %>%
  mutate(id_j = as.integer(id_j), metric = "Similarity")

Dbar_long <- as_tibble(Dbar, .name_repair = "minimal") %>%
  mutate(id_i = row_number()) %>%
  pivot_longer(cols = -id_i, names_to = "id_j", values_to = "val") %>%
  mutate(id_j = as.integer(id_j), metric = "Dissimilarity")

plot_long <- bind_rows(Sbar_long, Dbar_long)

# curv blocks: 1–9, 10–18, 19–27
curv_blocks <- tibble(
  curv  = c(0, 1, 2),
  start = c(1, 10, 19),
  end   = c(9, 18, 27),
  mid   = c(5, 14, 23),
  label = c("curv 0", "curv 1", "curv 2")
)

# nv blocks inside each curv block: n1=(1–3), n2=(4–6), n3=(7–9) (offset by curv*9)
nv_blocks <- tidyr::expand_grid(curv = c(0,1,2), nv = c(1,2,3)) %>%
  mutate(
    start = (curv*9) + (nv-1)*3 + 1,
    end   = start + 2,
    mid   = (start + end)/2,
    label = paste0("n", nv)
  )

# angle labels (per nv block): columns are [45,90,135] in each 3-wide chunk
ang_labs <- tibble(ang = c(45, 90, 135), offset = c(0,1,2), lab = c("a45","a90","a135"))
ang_positions <- nv_blocks %>%
  select(curv, nv, start) %>%
  tidyr::crossing(ang_labs) %>%
  mutate(x = start + offset, y = 0.4)  # y will be used for top annotations


# curv blocks: 1–9, 10–18, 19–27
curv_blocks <- tibble(
  curv  = c(0, 1, 2),
  start = c(1, 10, 19),
  end   = c(9, 18, 27),
  mid   = c(5, 14, 23),
  label = c("curv 0", "curv 1", "curv 2")
)

# nv blocks inside each curv block: n1=(1–3), n2=(4–6), n3=(7–9) (offset by curv*9)
nv_blocks <- tidyr::expand_grid(curv = c(0,1,2), nv = c(1,2,3)) %>%
  mutate(
    start = (curv*9) + (nv-1)*3 + 1,
    end   = start + 2,
    mid   = (start + end)/2,
    label = paste0("n", nv)
  )

# angle labels (per nv block): columns are [45,90,135] in each 3-wide chunk
ang_labs <- tibble(ang = c(45, 90, 135), offset = c(0,1,2), lab = c("a45","a90","a135"))
ang_positions <- nv_blocks %>%
  select(curv, nv, start) %>%
  tidyr::crossing(ang_labs) %>%
  mutate(x = start + offset, y = 0.4)  # y will be used for top annotations


plot_matrix <- function(M,
                        stim_lookup,
                        value_name = "Similarity",
                        title = NULL,
                        limits = c(0, 1),
                        na_value = "grey90",
                        palette = "plasma",
                        show_curv_labels = TRUE,
                        show_nv_labels = TRUE,
                        show_ang_right = TRUE,
                        base_size = 14) {

  stopifnot(is.matrix(M))
  stopifnot(all(c("curv","nv","ang","stim_id") %in% names(stim_lookup)))

  library(dplyr)
  library(tidyr)
  library(ggplot2)

  `%||%` <- function(x, y) if (is.null(x) || length(x) == 0) y else x

  # order: curv -> nv -> ang (matches your stim_lookup)
  stim_order <- stim_lookup %>%
    arrange(curv, nv, ang) %>%
    pull(stim_id)

  N <- length(stim_order)
  stopifnot(nrow(M) == N, ncol(M) == N)

  # long form matrix
  M_long <- as_tibble(M, .name_repair = "minimal") %>%
    mutate(id_i = row_number()) %>%
    pivot_longer(cols = -id_i, names_to = "id_j", values_to = "val") %>%
    mutate(
      id_i = as.integer(id_i),
      id_j = as.integer(id_j),
      id_i = factor(id_i, levels = stim_order),
      id_j = factor(id_j, levels = stim_order)
    )

  # blocks (hard-coded for 27 = 3*3*3; matches your setup)
  curv_blocks <- tibble(
    curv  = c(0, 1, 2),
    start = c(1, 10, 19),
    end   = c(9, 18, 27),
    mid   = c(5, 14, 23),
    label = c("curv 0", "curv 1", "curv 2")
  )

  nv_blocks <- tidyr::expand_grid(
    curv = c(0, 1, 2),
    nv   = c("n1", "n2", "n3")
  ) %>%
    mutate(
      start_local = case_when(nv == "n1" ~ 1L,
                              nv == "n2" ~ 4L,
                              nv == "n3" ~ 7L),
      end_local   = start_local + 2L,
      start = start_local + curv*9,
      end   = end_local   + curv*9,
      mid   = (start + end)/2
    )

  # RIGHT-SIDE ANG LABELS: one per row, derived from stim_lookup order
  # (we label rows, so we use the row ordering: arrange(curv, nv, ang))
  ang_right <- stim_lookup %>%
    arrange(curv, nv, ang) %>%
    transmute(
      y = stim_id,                 # 1..27 in your lookup
      label = paste0("a", ang)
    )

  # palette
  fill_scale <- scale_fill_viridis_c(
    option = if (palette == "plasma") "plasma" else "viridis",
    direction = -1,
    limits = limits,
    name = value_name,
    na.value = na_value
  )

  # base heatmap (square)
  p <- ggplot(M_long, aes(x = id_i, y = id_j, fill = val)) +
    geom_tile() +
    fill_scale +
    coord_equal(
      clip = "off",
      xlim = c(0.5, N + 0.5),
      ylim = c(0.5, N + 0.5)
    ) +
    theme_minimal(base_size = base_size) +
    theme(
      panel.grid  = element_blank(),
      axis.text   = element_blank(),
      axis.ticks  = element_blank(),
      plot.margin = margin(20, 130, 70, 90),  # extra right margin for ang labels
      aspect.ratio = 1
    ) +
    labs(x = NULL, y = NULL, title = title %||% "")

  # separators: curv (thick)
  for (x in c(9.5, 18.5)) {
    p <- p +
      geom_vline(xintercept = x, linewidth = 0.9) +
      geom_hline(yintercept = x, linewidth = 0.9)
  }

  # separators: nv within each curv (medium)
  for (x in c(3.5, 6.5, 12.5, 15.5, 21.5, 24.5)) {
    p <- p +
      geom_vline(xintercept = x, linewidth = 0.45, alpha = 0.9) +
      geom_hline(yintercept = x, linewidth = 0.45, alpha = 0.9)
  }

  # separators: ang within each nv (thin)
  thin_lines <- c(
    1.5,2.5,4.5,5.5,7.5,8.5,
    10.5,11.5,13.5,14.5,16.5,17.5,
    19.5,20.5,22.5,23.5,25.5,26.5
  )
  for (x in thin_lines) {
    p <- p +
      geom_vline(xintercept = x, linewidth = 0.22, alpha = 0.55) +
      geom_hline(yintercept = x, linewidth = 0.22, alpha = 0.55)
  }

  # curv labels (bottom)
  if (show_curv_labels) {
    p <- p +
      geom_text(
        data = curv_blocks,
        aes(x = mid, y = 0, label = label),
        inherit.aes = FALSE,
        vjust = 1.5,
        size = 5
      )
  }

  # nv labels (left)
  if (show_nv_labels) {
    p <- p +
      geom_text(
        data = nv_blocks,
        aes(x = 0, y = mid, label = nv),
        inherit.aes = FALSE,
        hjust = 1.4,
        size = 4.2,
        fontface = "bold"
      )
  }

  # ang labels (right) — one per row, small font
  if (show_ang_right) {
    p <- p +
      geom_text(
        data = ang_right,
        aes(x = N + 1.2, y = y, label = label),
        inherit.aes = FALSE,
        hjust = 0,
        size = 3.0,
        alpha = 0.9
      )
  }

  p
}


# Similarity heatmap (yellow -> dark blue look)
pS <- plot_matrix(
  M = Sbar,
  stim_lookup = stim_lookup,
  value_name = "Similarity",
  title = "Average similarity matrix (S̄)",
  limits = c(0, 1),
  palette = "plasma",
  show_ang_right = TRUE
)
pS

Dbar <- 1 - Sbar

pD <- plot_matrix(
  M = Dbar,
  stim_lookup = stim_lookup,
  value_name = "Dissimilarity",
  title = "Average dissimilarity matrix (D̄)",
  limits = c(0, 1),
  palette = "plasma",
  show_ang_right = TRUE
)

pD

```


```{r}

library(dplyr)
library(ggplot2)
library(smacof)
library(tibble)

# weights
W <- ifelse(is.na(Dbar), 0, 1)

# replace NA with 0
D_smacof <- Dbar
D_smacof[is.na(D_smacof)] <- 0

# Ordinal MDS (2D)
mds_ordinal <- mds(
  delta = D_smacof,
  weightmat = W,
  type = "ordinal",
  ndim = 2)

mds_ordinal$stress

# extract coordinates, join data about stimulus
mds_coords_ord <- as.data.frame(mds_ordinal$conf) %>%
  rownames_to_column("stim_id") %>%
  mutate(stim_id = as.integer(stim_id)) %>%
  left_join(stim_lookup, by = "stim_id")

head(mds_coords_ord)


```



```{r}
ggplot(mds_coords_ord, aes(x = D1, y = D2, label = stim_id)) +
  geom_point() +
  geom_text(vjust = -0.6, size = 3) +
  theme_minimal(base_size = 14) +
  labs(title = "Ordinal MDS (from D̄)", x = "Dim 1", y = "Dim 2")

# colored by curv
ggplot(mds_coords_ord, aes(x = D1, y = D2)) +
  geom_point(aes(color = factor(curv)), size = 3) +
  geom_text(aes(label = stim_id), vjust = -0.8, size = 3, show.legend = FALSE) +
  theme_minimal(base_size = 14) +
  labs(title = "Ordinal MDS (D̄), colored by curv", x = "Dim 1", y = "Dim 2", color = "curv")

# colored by nv
ggplot(mds_coords_ord, aes(x = D1, y = D2)) +
  geom_point(aes(color = factor(nv)), size = 3) +
  geom_text(aes(label = stim_id), vjust = -0.8, size = 3, show.legend = FALSE) +
  theme_minimal(base_size = 14) +
  labs(title = "Ordinal MDS (D̄), colored by nv", x = "Dim 1", y = "Dim 2", color = "nv")

# colored by ang
ggplot(mds_coords_ord, aes(x = D1, y = D2)) +
  geom_point(aes(color = factor(ang)), size = 3) +
  geom_text(aes(label = stim_id), vjust = -0.8, size = 3, show.legend = FALSE) +
  theme_minimal(base_size = 14) +
  labs(title = "Ordinal MDS (D̄), colored by ang", x = "Dim 1", y = "Dim 2", color = "ang")

mds_coords_ord %>%
  select(D1, D2, curv, nv, ang) %>%
  mutate(across(c(curv, nv, ang), as.numeric)) %>%
  cor(use = "complete.obs")

```


```{r}
# Ideal Matrices 

ideal_curv <- matrix(0, nrow = 27, ncol = 27)
ideal_nv <- matrix(0, 27, 27)
ideal_ang <- matrix(0, 27, 27)
ideal_additive <- matrix(0, 27, 27)

for (i in 1:27) {
  for (j in 1:27) {
    ideal_curv[i, j] <- as.integer(stim_lookup$curv[i] != stim_lookup$curv[j])
  }
}

for (i in 1:27) {
  for (j in 1:27) {
    ideal_nv[i, j] <- as.integer(stim_lookup$nv[i] != stim_lookup$nv[j])
  }
}

for (i in 1:27) {
  for (j in 1:27) {
    ideal_ang[i, j] <- as.integer(stim_lookup$ang[i] != stim_lookup$ang[j])
  }
}

for (i in 1:27) {
  for (j in 1:27) {
    ideal_additive[i, j] <-
      (stim_lookup$curv[i] != stim_lookup$curv[j]) +
      (stim_lookup$nv[i]   != stim_lookup$nv[j]) +
      (stim_lookup$ang[i]  != stim_lookup$ang[j])
  }
}

diag(ideal_curv) <- 0
diag(ideal_nv) <- 0
diag(ideal_ang) <- 0
diag(ideal_additive) <- 0


```

```{r}
n_items <- nrow(stim_lookup)

ideal_curv <- matrix(0, n_items, n_items)
ideal_nv   <- matrix(0, n_items, n_items)
ideal_ang  <- matrix(0, n_items, n_items)
ideal_add  <- matrix(0, n_items, n_items)

for (i in 1:n_items) {
  for (j in 1:n_items) {
    ideal_curv[i, j] <- as.integer(stim_lookup$curv[i] != stim_lookup$curv[j])
    ideal_nv[i, j]   <- as.integer(stim_lookup$nv[i]   != stim_lookup$nv[j])
    ideal_ang[i, j]  <- as.integer(stim_lookup$ang[i]  != stim_lookup$ang[j])

    ideal_add[i, j]  <-
      (stim_lookup$curv[i] != stim_lookup$curv[j]) +
      (stim_lookup$nv[i]   != stim_lookup$nv[j]) +
      (stim_lookup$ang[i]  != stim_lookup$ang[j])
  }
}

diag(ideal_curv) <- 0
diag(ideal_nv)   <- 0
diag(ideal_ang)  <- 0
diag(ideal_add)  <- 0

ideal_add01 <- ideal_add / 3
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

stim_order <- stim_lookup %>% arrange(curv, nv, ang) %>% pull(stim_id)

plot_mat <- function(M, title, lims = c(0,1)) {
  as_tibble(M, .name_repair = "minimal") %>%
    mutate(id_i = row_number()) %>%
    pivot_longer(cols = -id_i, names_to = "id_j", values_to = "D") %>%
    mutate(
      id_i = factor(as.integer(id_i), levels = stim_order),
      id_j = factor(as.integer(id_j), levels = stim_order)
    ) %>%
    ggplot(aes(id_i, id_j, fill = D)) +
    geom_tile() +
    scale_fill_viridis_c(option = "viridis", direction = -1,
                         limits = lims, name = "Dissimilarity") +
    coord_equal() +
    theme_minimal(base_size = 13) +
    theme(
      panel.grid = element_blank(),
      axis.text  = element_blank(),
      axis.ticks = element_blank()
    ) +
    labs(x = NULL, y = NULL, title = title)
}

colnames(Dbar) <- seq_len(ncol(Dbar))
colnames(ideal_curv) <- seq_len(ncol(ideal_curv))
colnames(ideal_nv)   <- seq_len(ncol(ideal_nv))
colnames(ideal_ang)  <- seq_len(ncol(ideal_ang))
colnames(ideal_add01) <- seq_len(ncol(ideal_add01))

p_emp   <- plot_mat(Dbar,        "Empirical D̄ (group average)", lims = c(0,1))
p_curv  <- plot_mat(ideal_curv,  "Ideal D, curv",          lims = c(0,1))
p_nv    <- plot_mat(ideal_nv,    "Ideal D, nv",            lims = c(0,1))
p_ang   <- plot_mat(ideal_ang,   "Ideal D, angle",         lims = c(0,1))
p_add   <- plot_mat(ideal_add01, "Ideal D, all",        lims = c(0,1))

p_emp
p_curv
p_nv
p_ang
p_add
```
```{r}
# Pearson correlation between ideal matrices and empirical matrix. 

get_rdm_vec <- function(M) {
  M[upper.tri(M, diag = FALSE)]
}

v_emp   <- get_rdm_vec(Dbar)
v_curv  <- get_rdm_vec(ideal_curv)
v_nv    <- get_rdm_vec(ideal_nv)
v_ang   <- get_rdm_vec(ideal_ang)
v_add   <- get_rdm_vec(ideal_add01)


ok <- !is.na(v_emp)
cor(v_emp[ok], v_curv[ok], method = "pearson")
cor(v_emp[ok], v_nv[ok],   method = "pearson")
cor(v_emp[ok], v_ang[ok],  method = "pearson")
cor(v_emp[ok], v_add[ok],  method = "pearson")
print('----------')
cor(v_emp[ok], v_curv[ok], method = "spearman")
cor(v_emp[ok], v_nv[ok],   method = "spearman")
cor(v_emp[ok], v_ang[ok],  method = "spearman")
cor(v_emp[ok], v_add[ok],  method = "spearman")
# Do stimulus pairs that are more dissimilar under the ideal model tend to be ranked as more dissimilar empirically?
```


```{r}
# Consistency between participants
# Split half correlation
# Uses S_subj -> builds half Dbar -> correlates with ideals

set.seed(1)

get_rdm_vec <- function(M) M[upper.tri(M, diag = FALSE)]

# helper function to build a 27x27 D matrix from a table of (low, high, S_mean)
make_D_from_S <- function(S_tbl, n_items = 27) {
  Sbar <- matrix(NA_real_, nrow = n_items, ncol = n_items)
  for (k in seq_len(nrow(S_tbl))) {
    i <- S_tbl$low[k]
    j <- S_tbl$high[k]
    Sbar[i, j] <- S_tbl$S_mean[k]
    Sbar[j, i] <- S_tbl$S_mean[k]
  }
  diag(Sbar) <- 1
  Dbar <- 1 - Sbar
  diag(Dbar) <- 0
  Dbar
}

# ideal vectors
v_curv <- get_rdm_vec(ideal_curv)
v_nv   <- get_rdm_vec(ideal_nv)
v_ang  <- get_rdm_vec(ideal_ang)
v_add  <- get_rdm_vec(ideal_add01)

subjs <- sort(unique(S_subj$subj_id))
n_subj <- length(subjs)
stopifnot(n_subj >= 4)  # sanity

n_iter <- 2000

out <- data.frame(
  iter = 1:n_iter,
  n1 = integer(n_iter),
  n2 = integer(n_iter),
  curv_h1 = NA_real_, curv_h2 = NA_real_,
  nv_h1   = NA_real_, nv_h2   = NA_real_,
  ang_h1  = NA_real_, ang_h2  = NA_real_,
  add_h1  = NA_real_, add_h2  = NA_real_
)

for (it in 1:n_iter) {

  idx <- sample(subjs)
  n1  <- floor(n_subj / 2)
  g1  <- idx[1:n1]
  g2  <- idx[(n1 + 1):n_subj]

  out$n1[it] <- length(g1)
  out$n2[it] <- length(g2)

  # half-average S for each pair (low, high)
  S1 <- S_subj %>%
    dplyr::filter(subj_id %in% g1) %>%
    dplyr::group_by(low, high) %>%
    dplyr::summarise(S_mean = mean(S, na.rm = TRUE), .groups = "drop")

  S2 <- S_subj %>%
    dplyr::filter(subj_id %in% g2) %>%
    dplyr::group_by(low, high) %>%
    dplyr::summarise(S_mean = mean(S, na.rm = TRUE), .groups = "drop")

  D1 <- make_D_from_S(S1, n_items = 27)
  D2 <- make_D_from_S(S2, n_items = 27)

  v1 <- get_rdm_vec(D1)
  v2 <- get_rdm_vec(D2)

  #
  ok <- !is.na(v1) & !is.na(v2)

  out$curv_h1[it] <- cor(v1[ok], v_curv[ok], method = "spearman")
  out$nv_h1[it]   <- cor(v1[ok], v_nv[ok],   method = "spearman")
  out$ang_h1[it]  <- cor(v1[ok], v_ang[ok],  method = "spearman")
  out$add_h1[it]  <- cor(v1[ok], v_add[ok],  method = "spearman")

  out$curv_h2[it] <- cor(v2[ok], v_curv[ok], method = "spearman")
  out$nv_h2[it]   <- cor(v2[ok], v_nv[ok],   method = "spearman")
  out$ang_h2[it]  <- cor(v2[ok], v_ang[ok],  method = "spearman")
  out$add_h2[it]  <- cor(v2[ok], v_add[ok],  method = "spearman")
}

# single number for stability per model (correlation of halffits across the iterations)
stability <- c(
  curv = cor(out$curv_h1, out$curv_h2, use = "complete.obs"),
  nv   = cor(out$nv_h1,   out$nv_h2,   use = "complete.obs"),
  ang  = cor(out$ang_h1,  out$ang_h2,  use = "complete.obs"),
  add  = cor(out$add_h1,  out$add_h2,  use = "complete.obs")
)
print(round(stability, 3))

# out dataframe
# h1: the Spearman correlation between the half-1 empirical dissimilarity vector and the ideal model vector
# h2: the same correlation but for half-2

```


Stability of N subjects analiz

1. RDM stability vs N (subsample vs full 17-subject reference)

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

get_rdm_vec <- function(M) M[upper.tri(M, diag = FALSE)]

make_D_from_S <- function(S_tbl, n_items = 27) {
  Sbar <- matrix(NA_real_, nrow = n_items, ncol = n_items)
  for (k in seq_len(nrow(S_tbl))) {
    i <- S_tbl$low[k]
    j <- S_tbl$high[k]
    Sbar[i, j] <- S_tbl$S_mean[k]
    Sbar[j, i] <- S_tbl$S_mean[k]
  }
  diag(Sbar) <- 1
  Dbar <- 1 - Sbar
  diag(Dbar) <- 0
  Dbar
}


# full 17 subject RDM vector 
n_items <- 27

S_full_tbl <- S_subj %>%
  group_by(low, high) %>%
  summarise(S_mean = mean(S, na.rm = TRUE), .groups = "drop")

D_full <- make_D_from_S(S_full_tbl, n_items = n_items)
v_full <- get_rdm_vec(D_full)


```

```{r}
# without replacement sampling of subjects
set.seed(1)

subjs <- sort(unique(S_subj$subj_id))
n_subj <- length(subjs)

N_grid <- c(2,4,6,8,10,12,14,16)  
n_iter <- 1000

stab <- tidyr::expand_grid(N = N_grid, iter = 1:n_iter) %>%
  mutate(r = NA_real_)

for (k in seq_len(nrow(stab))) {
  N  <- stab$N[k]
  ss <- sample(subjs, size = N, replace = FALSE)

  S_N_tbl <- S_subj %>%
    filter(subj_id %in% ss) %>%
    group_by(low, high) %>%
    summarise(S_mean = mean(S, na.rm = TRUE), .groups = "drop")

  D_N <- make_D_from_S(S_N_tbl, n_items = n_items)
  v_N <- get_rdm_vec(D_N)

  ok <- !is.na(v_N) & !is.na(v_full)

  # spearman correlation
  stab$r[k] <- cor(v_N[ok], v_full[ok], method = "spearman")
}
```

```{r}
stab_sum <- stab %>%
  group_by(N) %>%
  summarise(
    r_mean = mean(r, na.rm = TRUE),
    r_med  = median(r, na.rm = TRUE),
    r_lo   = quantile(r, 0.025, na.rm = TRUE),
    r_hi   = quantile(r, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(stab_sum, aes(x = N, y = r_mean)) +
  geom_ribbon(aes(ymin = r_lo, ymax = r_hi), alpha = 0.2) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = N_grid) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal(base_size = 14) +
  labs(
    x = "N subjects in subsample",
    y = "Spearman corr")
```

2) split-half stability vs N
For each iteration, samples 2N subjects without replacement, splits into half1 and half2 each with N. builds Dissimilarity matrix for each half. 
computes Spearman correlation between the two RDM vectors. 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

set.seed(1)

subjs <- sort(unique(S_subj$subj_id))
n_subj <- length(subjs)
n_items <- 27

# 17 subjects, so max N for split-half is 8
N_grid <- 2:floor(n_subj/2)   # c(2,4,6,8) 
n_iter <- 2000

stab_sh <- tidyr::expand_grid(N = N_grid, iter = 1:n_iter) %>%
  mutate(r = NA_real_)

for (k in seq_len(nrow(stab_sh))) {
  N <- stab_sh$N[k]

  # sample 2N subjects, then split into two halves of size N
  ss <- sample(subjs, size = 2*N, replace = FALSE)
  g1 <- ss[1:N]
  g2 <- ss[(N+1):(2*N)]

  S1 <- S_subj %>%
    filter(subj_id %in% g1) %>%
    group_by(low, high) %>%
    summarise(S_mean = mean(S, na.rm = TRUE), .groups = "drop")

  S2 <- S_subj %>%
    filter(subj_id %in% g2) %>%
    group_by(low, high) %>%
    summarise(S_mean = mean(S, na.rm = TRUE), .groups = "drop")

  # convert to dissimilarity matrices and vectors
  D1 <- make_D_from_S(S1, n_items = n_items)
  D2 <- make_D_from_S(S2, n_items = n_items)

  v1 <- get_rdm_vec(D1)
  v2 <- get_rdm_vec(D2)

  ok <- !is.na(v1) & !is.na(v2)

  stab_sh$r[k] <- cor(v1[ok], v2[ok], method = "spearman")
}

stab_sh_sum <- stab_sh %>%
  group_by(N) %>%
  summarise(
    r_mean = mean(r, na.rm = TRUE),
    r_med  = median(r, na.rm = TRUE),
    r_lo   = quantile(r, 0.025, na.rm = TRUE),
    r_hi   = quantile(r, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

# plot
ggplot(stab_sh_sum, aes(x = N, y = r_mean)) +
  geom_ribbon(aes(ymin = r_lo, ymax = r_hi), alpha = 0.2) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = N_grid) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal(base_size = 14) +
  labs(
    x = "N subjects per half",
    y = "Spearman corr",
    title = "Split-half RDM stability vs N"
  )
```

